# Week 6 - Analysis of longitudinal data

## Part 1 - Implement the analyses of Chapter 8 of MABS using the RATS data.

Read the data:

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(GGally)
library(ggplot2)
RATS <- read.csv(file = "~/IODS-project/data/RATS.csv", header = TRUE, row.names = 1)
str(RATS)
```

Factor the categorical variables:

```{r}
RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)
str(RATS)
```
```{r}
ggplot(RATS, aes(x = Time, y = weight, color=Group, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)")
```

```{r}
ggplot(RATS, aes(x = Time, y = weight, color=Group, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Weight (grams)") +
  facet_grid(. ~ Group, labeller = label_both)
```

The graph shows the weight change of 3 groups of rats during the study which has a duration of 60 days. Looking at the graph, we can observe that the tendency to gain weight is higher for group 2 and 3 as compared to group 1.

Let's standardize the variables

```{r}
library(ggplot2)
library(dplyr) 
library(tidyr)
library(lme4)
RATS_1 <- RATS %>%
  group_by(Time) %>%
  mutate(stdrats = (weight - mean(weight))/sd(weight) ) %>%
  ungroup()
glimpse(RATS_1)
```

Plot using the standardize data:

```{r}
ggplot(RATS_1, aes(x = Time, y = stdrats, color=ID, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Standardized Weight (grams)")
```

```{r}
ggplot(RATS_1, aes(x = Time, y = stdrats, color=Group, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "Standardized Weight (grams)") +
  facet_grid(. ~ Group, labeller = label_both)
```

The mean profiles suggest there is a huge difference between the 3 groups with respect to mean weight. 

Summary graph using means profiles:

```{r}
n <- RATS$Time %>% unique() %>% length()
RATS_2 <- RATS_1 %>%
  group_by(Group, WD) %>%
  summarise( mean = mean(weight), se = sd(weight)/sqrt(n) ) %>%
  ungroup()
ggplot(RATS_2, aes(x = WD, y = mean, linetype = "ID", shape = "ID")) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "Mean(Weight) +/- Se(Weight)")
```

This graph shows means profile for individual subjects of experiment. It can be used to compare different subjects based on their single measurements taken  as a mean from the measurements over the period of the study.

We can also check the impact of outliers in the study. 

```{r}
RATS_3 <- RATS_1 %>%
  filter(Time > 0) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(weight) ) %>%
  ungroup()
ggplot(RATS_3, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "blue") +
  scale_y_continuous(name = "mean(Weight)")
```

There are outliers in data which can be seen clearly here. The most significant outlier can be found in group 2 which we can remove for better analysis. 

```{r}
RATS_4 <- RATS_3 %>% filter(mean <570)
```
```{r}
oneway.test(mean~Group,data=RATS_4,var.equal=TRUE) 
```

Using the results of  anova test above, it can be 


## Part 2 - Implement the analyses of Chapter 9 of MABS using the BPRS data. 

Read the data:

```{r}
BPRS <- read.csv(file = "~/IODS-project/data/BPRS.csv", header = TRUE, row.names = 1)
str(BPRS)
```

Factor the categorical variables and converting weeks from factor to char. 

```{r}
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
BPRS$weeks <- as.character(BPRS$weeks)
str(BPRS)
```

Standardize the data:

```{r}
BPRS <- BPRS %>%
  group_by(week) %>%
  mutate(stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()
str(BPRS)
```

```{r}
ggpairs(BPRS[,c(1,4:5)])
```

Plotting the data:

```{r}
ggplot(BPRS, aes(x = week, y = bprs, color = subject)) +
  geom_line(aes(linetype = subject)) +
  scale_x_continuous(name = "Week", breaks = seq(0, 60, 10)) +
  scale_y_continuous(name = "bprs") +
  theme(legend.position = "top")
```

Plotting the data for both treatment types:

```{r}
ggplot(BPRS, aes(x = week, y = bprs, color=subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRS$bprs), max(BPRS$bprs))) +
  theme(legend.position = "top")
```



```{r}
ggplot(BPRS, aes(x = week, y = bprs, color = treatment)) +
  geom_line(aes(linetype = treatment)) +
  scale_x_continuous(name = "week", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "bprs") +
  theme(legend.position = "top")
```

Using multiple regression model:

```{r}
BPRS_1 <- lm(bprs ~ week + treatment, data = BPRS)
summary(BPRS_1)
```

We can see that week number (time) is more significant than treatment type. 

Using the random intercept model:

```{r}

BPRS_2 <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRS, REML = FALSE)
summary(BPRS_2)
```

The result shows that estimated variance of the random effects is really high. There are visible changes in values of treatment2 but are not significant. 

Using the random intercept and random slope model:

```{r}
BPRS_3 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRS, REML = FALSE)
summary(BPRS_3)
```

Using anova to compare models:

```{r}
anova(BPRS_3, BPRS_2)
```



